### 2015-06-09 LOI/inc2.tex
Fixed a typo.

\underline{\textbf{Objectives:}} Exascale scientific discovery will be
severely bottlenecked without sufficient new research into managing and
storing the large amounts of data that will be produced during the
simulation, and analyzed for months afterwards.  Our goal in this project is
to address the associated I/O and storage challenges in the context of
current and emerging storage landscapes, and expedite insights into mission
critical scientific processes. To that end, we will build on the
capabilities offered by ADIOS and DataSpaces at ORNL that provide I/O abstractions
and services, the Sirocco peer-to-peer file system at
Sandia and the object storage and annotation expertise of UC Santa Cruz, to
explore application and multi-tier storage aware data management
solutions.    This project
brings together a  team with strong expertise in I/O middleware (ORNL), file system (SNL, UCSC) and storage (UCSC), and connects and
coordinates these key storage components in a seamless fashion.

@@

% Our proposed I/O and storage system prototype will enable
% highly flexible data access modes that are common in scientific applications
% and associated analysis.
%
%\hasan{Can we actually take some lower accuracy representation of data (low
%  granularity MR for example, and compute more stuff to provide a higher
%  accuracy representation? I am thinking of the work with Varis Carey with
%  storing boundary state instead of full volume and then using those
%  boundaries to fill in blanks over temporally changing data}
%
The proposed work will address the following technical areas;
\\\textbf{First}, the storage system will offer the ability to directly store
different versions of a single variable to different
levels of the storage hierarchy and allow each version to be processed 
differently. For example, full data can be queued to tape while
a highly compressed version intended for high-level analytical views can be
stored in non-volatile memory (NVM) or even RAM within the storage
hierarchy. 
The key challenge here would be defining and maintaining the metadata
connecting differrent object quality and utility levels, 
The storage system will support the metadata and plug-ins required to support
this data storage approach.   
%data re-computation is allowed to re-generate data with
%customizable accuracy, if the associated computing and I/O cost is lower
%than retrieving the target data from a {\em slow} tier.
%
\textbf{Second}, new storage access APIs will be developed that incorporates the
notion of time and data quality for requested data. The storage system,
through the knowledge of both data quality stored in different storage
hierarchy tiers and the approximate time to retrieve data from the different
tiers, can manage (with the user in the loop) the tradeoff between error bound and the time to
retrieve data. This storage system support will be executed and managed through our middleware, 
insulating the user as much as possible from the new APIs. The storage system
will also support plug-ins, for example, to potentially decompress or expand data to the
original size. Both lossless and lossy with error bounds style storage is
assumed. The time factor must take into account how long this operation takes
for servicing the user request. 
%exploratory analysis operations frequently entail an
%overall data set view followed by targeted data exploration on extracted
%features.  In particular, the {\em overview} access mode gives a quick but
%approximate data view that can guide feature selection offering rapid
%coarse-grained data exploration without requiring loading the high-accuracy
%data from storage. Based on the granularity requested, the accuracy and size
%of the data returned can be adjusted. In the most extreme setting, the
%original data can be retrieved at the time cost of moving the potentially
%huge data quantity.
%
\textbf{Third}, the storage system will offer annotations within the metadata
and support for both predictive and reactive data placement and migration.
While past access patterns may not indicate future access because the
simulation details may have changed, we are focused on scalability where
subsequent runs are larger in scale as the simulation prepares for a capability run.  By
learning from the output and access patterns during this run sequence, we can
accurately decide how to place and organize data for the critical capability
runs. The reactive mechanisms will consider the space and performance
characteristics, as well as the requested data fidelity, to determine where and
how to pull and store data sets. The challenge here is to utilize this
metadata to manage the placement and access of data object based on the
defined constraints as well as storage system state.

>>>>

\underline{\textbf{Objectives:}} Exascale scientific discovery will be
severely bottlenecked without sufficient new research into managing and
storing the large amounts of data that will be produced during the
simulation, and analyzed for months afterwards.  Our goal in this project is
to address the associated I/O and storage challenges in the context of
current and emerging storage landscapes, and expedite insights into mission
critical scientific processes. To that end, we will build on the
capabilities offered by ADIOS and DataSpaces at ORNL that provide I/O abstractions
and services, the Sirocco peer-to-peer file system at
Sandia and the object storage and annotation expertise of UC Santa Cruz, to
explore application and multi-tier storage aware data management
solutions.    This project
brings together a  team with strong expertise in I/O middleware (ORNL, Rutgers), file system (SNL, UCSC) and storage (UCSC), and connects and
coordinates these key storage components in a seamless fashion.

@@

% Our proposed I/O and storage system prototype will enable
% highly flexible data access modes that are common in scientific applications
% and associated analysis.
%
%\hasan{Can we actually take some lower accuracy representation of data (low
%  granularity MR for example, and compute more stuff to provide a higher
%  accuracy representation? I am thinking of the work with Varis Carey with
%  storing boundary state instead of full volume and then using those
%  boundaries to fill in blanks over temporally changing data}
%
The proposed work will address the following technical areas;
\\\textbf{First}, the storage system will offer the ability to directly store
different versions of a single variable to different
levels of the storage hierarchy and allow each version to be processed 
differently. For example, full data can be queued to tape while
a highly compressed version intended for high-level analytical views can be
stored in non-volatile memory (NVM) or even RAM within the storage
hierarchy. 
The key challenge here would be defining and maintaining the metadata
connecting different object quality and utility levels, 
The storage system will support the metadata and plug-ins required to support
this data storage approach.   
%data re-computation is allowed to re-generate data with
%customizable accuracy, if the associated computing and I/O cost is lower
%than retrieving the target data from a {\em slow} tier.
%
\textbf{Second}, new storage access APIs will be developed that incorporates the
notion of time and data quality for requested data. The storage system,
through the knowledge of both data quality stored in different storage
hierarchy tiers and the approximate time to retrieve data from the different
tiers, can manage (with the user in the loop) the tradeoff between error bound and the time to
retrieve data. This storage system support will be executed and managed through our middleware, 
insulating the user as much as possible from the new APIs. The storage system
will also support plug-ins, for example, to potentially decompress or expand data to the
original size. Both lossless and lossy with error bounds style storage is
assumed. The time factor must take into account how long this operation takes
for servicing the user request. 
%exploratory analysis operations frequently entail an
%overall data set view followed by targeted data exploration on extracted
%features.  In particular, the {\em overview} access mode gives a quick but
%approximate data view that can guide feature selection offering rapid
%coarse-grained data exploration without requiring loading the high-accuracy
%data from storage. Based on the granularity requested, the accuracy and size
%of the data returned can be adjusted. In the most extreme setting, the
%original data can be retrieved at the time cost of moving the potentially
%huge data quantity.
%
\textbf{Third}, the storage system will offer annotations within the metadata
and support for both predictive and reactive data placement and migration.
While past access patterns may not indicate future access because the
simulation details may have changed, we are focused on scalability where
subsequent runs are larger in scale as the simulation prepares for a capability run.  By
learning from the output and access patterns during this run sequence, we can
accurately decide how to place and organize data for the critical capability
runs. The reactive mechanisms will consider the space and performance
characteristics, as well as the requested data fidelity, to determine where and
how to pull and store data sets. The challenge here is to utilize this
metadata to manage the placement and access of data object based on the
defined constraints as well as storage system state.
